---
title: "Your First Project Write-Up"
description: "A template for writing up a data science project with R code."
date: "2026-03-18"
categories: [R, visualization, statistics]
image: ""   # optional: path to a preview image
---

## Overview

Start with the *why*. What question were you trying to answer? Who would care about this?

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
# library(tidymodels)  # uncomment if modeling
```

## The Data

Describe your data source. Where did it come from? What does each row represent?

```{r}
#| label: load-data

# Replace with your actual data loading
# data <- read_csv("data/your-file.csv")

# Example with built-in data:
glimpse(mtcars)
```

## Exploration

What did you find when you first looked at the data?

```{r}
#| label: eda-plot
#| fig-cap: "A descriptive caption for your plot"

mtcars |>
  ggplot(aes(x = wt, y = mpg, color = factor(cyl))) +
  geom_point(size = 3, alpha = 0.8) +
  scale_color_manual(values = c("#1a3a5c", "#2e86ab", "#a8d8ea")) +
  labs(
    title = "Your Plot Title",
    x = "X Axis Label",
    y = "Y Axis Label",
    color = "Group"
  ) +
  theme_minimal(base_size = 13)
```

## Analysis

Walk through your methodology here. Code-fold is on by default so readers can toggle code.

```{r}
#| label: model

# Your modeling code here
summary(lm(mpg ~ wt + cyl, data = mtcars))
```

## Results

What did you find? Lead with the finding, then support it with the output above.

## Takeaways

- What worked well?
- What would you do differently?
- What's the next question this raises?
